Geometric models reveal the hidden structure of conceptual knowledge

### 2023-01-25 paper/main.tex
backing up progress before stopping for a meeting...

@@

\subsubsection*{Testing the ``smoothness'' of knowledge}

For each participant, for each quiz, for each of $c$ correctly answered
question $q$ in turn, we sorted all of the other correctly answered questions
in descending order of their topic vectors' correlations with $q$'s topic
vector. We repeated this for all correctly answered questions to obtain a $c
\times (c - 1)$ matrix of correlations, where the rows corresponded to held-out
questions and the columns corresponded to the most similar (leftmost column),
second most similar (second column), other correctly answered questions. Note
that each correlation for each pair of questions ($a$ and $b$) appears exactly
twice in this matrix: once when $a$ is the held-out question, and once when $b$
is the held-out question. We applied the Fisher $z$-transformation to these
correlations, averaged down the rows, and then applied the inverse Fisher
$z$-transformation to obtain a $1 \times (c - 1)$ array of average correlations
for the participant. We repeated this process across all $n$ participants,
yielding an $n \times 14$ matrix (15 is the maximum number of questions that
could have been answered correctly on any single 15-question quiz, so at most
14 correctly answered questions could be compared to a held-out correctly
answered question). Note that some of the entries, for some of the
participants, were empty (e.g., if a given participant answered at least one
question incorrectly, they would never have 14 other correctly answered
questions to compare to a held-out correctly answered question). If a
participant answered \textit{no} questions correctly on a given quiz, their
entire row would be empty.

>>>>

For each participant, for each quiz, for each of $c$ correctly answered
question $q$ in turn, we sorted all of the other correctly answered questions
in descending order of their topic vectors' correlations with $q$'s topic
vector. We repeated this for all correctly answered questions to obtain a $c
\times (c - 1)$ matrix of correlations, where the rows corresponded to held-out
questions and the columns corresponded to the most similar (leftmost column),
second most similar (second column), other correctly answered questions. Each
correlation for each pair of questions ($a$ and $b$) appears exactly twice in
this matrix: once when $a$ is the held-out question, and once when $b$ is the
held-out question. We applied the Fisher $z$-transformation to these
correlations, averaged down the rows, and then applied the inverse Fisher
$z$-transformation to obtain a $1 \times (c - 1)$ array of average correlations
for the participant. We repeated this process across all $n$ participants,
yielding an $n \times 14$ matrix (15 is the maximum number of questions that
could have been answered correctly on any single 15-question quiz, so at most
14 correctly answered questions could be compared to a held-out correctly
answered question). Note that some of the entries, for some of the
participants, were empty (e.g., if a given participant answered at least one
question incorrectly, they would never have 14 other correctly answered
questions to compare to a held-out correctly answered question). If a
participant answered \textit{no} questions correctly on a given quiz, their
entire row would be empty.

The correlation plots for both curves, in all three top panels of
Figure~\ref{fig:predictions}, confirm one aspect of our ``smoothness''
assumption. Specifically, given the topic vector of a correctly answered
question, other correctly answered questions also tend to have similar topic
vectors...

\textbf{JRM NOTE 1: need to think about these results more; stopping here
temporarily to back up progress...}

\textbf{JRM NOTE 2: consider moving ``estimating held-out conceptual
knowledge'' from methods to here.}

Knowledge estimates need not be limited to the content of the lectures. As
illustrated in Figure~\ref{fig:knowledge-maps}, our general approach to
estimating knowledge from a small number of quiz questions may be applied to
\textit{any} content, given its text embedding coordinate. To visualize how
knowledge ``spreads'' through text embedding space to content beyond the
lectures participants watched, we first fit a new topic model to the lectures'
sliding windows with $k = 100$ topics. We hoped that increasing the number of
topics from 15 to 100 might help us to generalize the knowledge predictions.
(Aside from increasing the number of topics from 15 to 100, all other
procedures and model parameters were carried over from the preceding analyses.)

@@
